NumPy---Array Programming in Python

### 2020-01-10 body.tex
Use array programming

Today, NumPy underpins almost every Python library that does scientific or
numerical computation including SciPy\cite{virtanen2019scipy},
matplotlib\cite{hunter2007matplotlib}, pandas\cite{mckinney-proc-scipy-2010},
scikit-learn\cite{pedregosa2011scikit}, and
scikit-image\cite{vanderwalt2014scikit}.
It consists of an $n$-dimensional array object along with utility functions
that operate on it.
Because of its inherent simplicity---being a pointer to memory with some
associated meta-data---the NumPy array is
the {\it de facto} exchange format for array data in Python.
The library has such widespread adoption that not only the array object but also its
{\it Application Programming Interface} (API) has become ubiquitous as
a language for array computation.

@@

NumPy was initially developed by students, faculty, and researchers in their
spare time to provide a modern array computation library for Python.
They were influenced by their experiences with powerful interactive programming
languages for scientific computing like Yorick \cite{munro1995using} as well
as commercial languages like IDL and Matlab.
Often they developed code to solve their own or their colleagues' problems.

>>>>

Today, NumPy underpins almost every Python library that does scientific or
numerical computation including SciPy\cite{virtanen2019scipy},
matplotlib\cite{hunter2007matplotlib}, pandas\cite{mckinney-proc-scipy-2010},
scikit-learn\cite{pedregosa2011scikit}, and
scikit-image\cite{vanderwalt2014scikit}.
It consists of an $n$-dimensional array object along with utility functions
that operate on it.
Because of its inherent simplicity---being a pointer to memory with some
associated meta-data---the NumPy array is
the {\it de facto} exchange format for array data in Python.
The library has such widespread adoption that not only the array object but also its
{\it Application Programming Interface} (API) has become ubiquitous as
a language for array programming.

@@

NumPy was initially developed by students, faculty, and researchers in their
spare time to provide a modern array programming library for Python.
They were influenced by their experiences with powerful interactive programming
languages for scientific computing like Yorick \cite{munro1995using} as well
as commercial languages like IDL and Matlab.
Often they developed code to solve their own or their colleagues' problems.

### 2020-01-10 summary.tex
Use array programming

% - 2-3 sentences of basic-level introduction to the field;
Array computation libraries provide powerful notation for accessing,
manipulating, and operating on data in vectors, matrices, and
higher-dimensional arrays \cite{iverson1980notation}.
% - a brief account of the background and rationale of the work;
NumPy is the fundamental array computation library for the Python ecosystem
\cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
It has been actively developed for over two decades and has more than 850
contributors, nearly 300,000 dependent repositories on GitHub, and millions of
downloads per year.
It plays an essential role in research analysis pipelines in fields as
diverse as physics, biology, astronomy, neuroscience, material science,
engineering, and chemistry.
For example, in astronomy, NumPy was fundamental to discovering gravitational
waves\cite{pycbc}, the first imaging of a black hole\cite{eht-imaging}, and
continues underpinning the future exploration of our
universe\cite{jenness2018lsst}.
It is one of the most commonly used machine learning tools among
enterprises \cite{451report2018}.
% - a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent);
Here we show how a small group of students and scientists created this
popular tool, explain the core ideas that helped make it it successful,
and highlight recent development to better serve new use cases
and future improvements.
% - 2-3 sentences putting the main findings into general context so it is
%   clear how the results described in the paper have moved the field forwards.
NumPy continues to evolve to meet the needs of its community, to remain
relevant in an ever-changing computing landscape, and to be a fundamental
pillar of support for the next decade of scientific computing, data science,
and machine learning in Python.

>>>>

% - 2-3 sentences of basic-level introduction to the field;
Array programming provides powerful notation for accessing,
manipulating, and operating on data in vectors, matrices, and
higher-dimensional arrays \cite{iverson1980notation}.
% - a brief account of the background and rationale of the work;
NumPy is the fundamental array programming library for the Python ecosystem
\cite{dubois2007guest,oliphant2007python,millman2011python,perez2011python}.
It has been actively developed for over two decades and has more than 850
contributors, nearly 300,000 dependent repositories on GitHub, and millions of
downloads per year.
It plays an essential role in research analysis pipelines in fields as
diverse as physics, biology, astronomy, neuroscience, material science,
engineering, and chemistry.
For example, in astronomy, NumPy was fundamental to discovering gravitational
waves\cite{pycbc}, the first imaging of a black hole\cite{eht-imaging}, and
continues underpinning the future exploration of our
universe\cite{jenness2018lsst}.
It is one of the most commonly used machine learning tools among
enterprises \cite{451report2018}.
% - a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent);
Here we show how a small group of students and scientists created this
popular tool, explain the core ideas that helped make it it successful,
and highlight recent development to better serve new use cases
and future improvements.
% - 2-3 sentences putting the main findings into general context so it is
%   clear how the results described in the paper have moved the field forwards.
NumPy continues to evolve to meet the needs of its community, to remain
relevant in an ever-changing computing landscape, and to be a fundamental
pillar of support for the next decade of scientific computing, data science,
and machine learning in Python.
